---
title: "EDA"
---

The data is only from 2018, however I am not confident enough in web-scrapping
to grab new data so I will use the 2018 set.

```{r directory}
knitr::opts_knit$set(root.dir = getwd())
```

```{r packages, message=FALSE, warning=FALSE}
source("packages.R")
for (i in seq_along(packages))
  library(packages[i], character.only = T)
rm(packages)
```

```{r data}
arabica_data <- fread("coffee-quality-database/data/arabica_data_cleaned.csv")
robusta_data <- fread("coffee-quality-database/data/robusta_data_cleaned.csv")
```

Having a look at the difference between arabica and robusta names
```{r names}
list(
  arabica_robusta = setdiff(names(arabica_data), names(robusta_data)),
  robust_arabica = setdiff(names(robusta_data), names(arabica_data))
)
```

Going to assume the following mapping
```{r names mapping}
setnames(
  robusta_data,
  old = setdiff(names(robusta_data), names(arabica_data)),
  new = c(
    "Aroma",
    "Acidity",
    "Sweetness",
    "Body",
    "Uniformity"
  )
)
```

Create species column and combine datasets
```{r species and combine}
combined_data <- rbind(robusta_data, arabica_data)
```

Look at summary data
```{r summary}
summary(combined_data)
```

V1 appears to be row number, can remove.
The differences in column names Camel.Case and snake_case annoys me 

#### Columns that seem important from the README.md in coffee-quality-database and from the classification challenge document.
* Aroma
* Flavor
* Aftertaste
* Acidity
* Body
* Balance
* Uniformity
* Cup Cleanliness
* Sweetness
* Moisture
* Defects
* Country of Origin
* Region
* Farm

#### Employee profiles
* Subu:     {indian, sweet, not too strong of flavour}
* Marcio:   {brazillian, regional, farm}
* Rafa:     same as Marcio
* James:    {soft, no armora, no flavour, very sweet}

```{r find Marcios profile from data}
combined_data[Farm.Name == "fazenda kaquend"]
```

https://soloespresso.net/coffee-cupping/
Used this website to understand the differences in coffee measures

* Marcio:     {Sweetness = 10, Flavor = 8.5, Acidity = 8, Body = 8, Aroma = 8.5, Balance = 8, Aftertaste = 8}
* Subu:       {Sweetness = 10, Flavor < 8}
* James:      {Aroma = 0, Flavor = 0, Sweetness = 10}

```{r summary of important variables}
summary(combined_data[, .(Sweetness, Flavor, Acidity, Body, Aroma, Balance, Aftertaste)])
```

They are all numeric variables except for country of origin, can use kmeans
or krpototypes if country is important

```{r countries}
sort(unique(combined_data$Country.of.Origin))
```

* Marcio:     {Sweetness = 10, Flavor = 8, Acidity = 8, Body = 8, Aroma = 8, Balance = 8, Aftertaste = 8, Country = 'Brazil'}
* Rafa:       {Sweetness = 10, Flavor = 8, Acidity = 8, Body = 8, Aroma = 8, Balance = 8, Aftertaste = 8, Country = 'Brazil'}
* Subu:       {Sweetness = 10, Flavor < 8,                                                                Country = 'India'}
* James:      {Sweetness = 10, Flavor = 0,                        Aroma = 0,                              Country = 'United States (Hawaii)'}

Assuming the coffee James and Subu drink coffee local to their home countries


```{r Flavor = 0}
combined_data[Flavor <= 0.1]
```
There's one data point full of 0's, so we will omit it as it's awful coffee
```{r}
combined_data <- combined_data[Flavor > 0.1]
```

```{r factor country}
# Create factor column
combined_data[, Country.of.Origin := as.factor(Country.of.Origin)]
```

```{r k-prototypes clustering, message=FALSE}
# Lambda test
lambda <- lambdaest(
  x = combined_data[,
                       .(
                         Sweetness,
                         Flavor,
                         Acidity,
                         Body,
                         Aroma,
                         Balance,
                         Aftertaste,
                         Country.of.Origin
                       )]
)

# K-prototypes clustering, find the best value of k
clusters <- validation_kproto(
  method = 'silhouette',
  data = combined_data[,
                       .(
                         Sweetness,
                         Flavor,
                         Acidity,
                         Body,
                         Aroma,
                         Balance,
                         Aftertaste,
                         Country.of.Origin
                       )],
  k = 2:floor(
    sqrt(
      nrow(
        combined_data
        )
      )
    ),
  lambda = lambda,
  verbose = F
)
clusters
```

3 clusters is optimal, however there are far too many coffees in these clusters
for Marcio to purchase.

```{r higher K}
# K-prototypes clustering, find the best value of k
clusters <- validation_kproto(
  method = 'silhouette',
  data = combined_data[,
                       .(
                         Sweetness,
                         Flavor,
                         Acidity,
                         Body,
                         Aroma,
                         Balance,
                         Aftertaste,
                         Country.of.Origin
                       )],
  k = 5:floor(
    sqrt(
      nrow(
        combined_data
        )
      )
    ),
  lambda = lambda,
  verbose = F
)
clusters
```
Still too many coffees, however there are now separate country modal clusters
for Hawaii and India

```{r try very high k}
# K-prototypes clustering, find the best value of k
clusters <- validation_kproto(
  method = 'silhouette',
  data = combined_data[,
                       .(
                         Sweetness,
                         Flavor,
                         Acidity,
                         Body,
                         Aroma,
                         Balance,
                         Aftertaste,
                         Country.of.Origin
                       )],
  k = 35:floor(
    sqrt(
      nrow(
        combined_data
        )
      )
    ),
  lambda = lambda,
  verbose = F
)
clusters
```

Data is more diverse now, zoom in on countries
```{r}
# Create cluster index
cluster_centres <- copy(clusters$kp_obj$centers)
cluster_centres[, cluster := 1:36]
cluster_centres[Country.of.Origin %in% c("India", "United States (Hawaii)", "Brazil")]
```
Find the coffees which relate to these clusters, I'm going to take the
sweetest Hawaiian one (31)
```{r coffees}
combined_data[, cluster := clusters$kp_obj$cluster]

chosen_clusters <- combined_data[cluster %in% c(21, 27, 31)]
```

Take a look at these coffees
```{r chosen_clusters view}
chosen_clusters
```

Remove all coffees with defects
```{r}
clean_clusters <- chosen_clusters[Category.One.Defects == 0 &
                                    Category.Two.Defects == 0]
```


Filter countries
```{r filter countries}
country_clusters <- clean_clusters[Country.of.Origin %in% c(
  "United States (Hawaii)",
  "Brazil",
  "India"
)]
```

```{r number of coffees}
table(country_clusters$Country.of.Origin %>% as.character)
```
check Marcio's farm
```{r farm check}
country_clusters[Farm.Name == "fazenda kaquend"]
```

It seem's like Marcio's farm has been omitted in the process, but he probably
doesn't need to be recommended his favourite coffee.

Now to find coffees from the list which match the profiles by creating a 
'taste distance'

* Marcio:     {Sweetness = 10, Flavor = 8, Acidity = 8, Body = 8, Aroma = 8, Balance = 8, Aftertaste = 8, Country = 'Brazil'}
* Rafa:       {Sweetness = 10, Flavor = 8, Acidity = 8, Body = 8, Aroma = 8, Balance = 8, Aftertaste = 8, Country = 'Brazil'}
* Subu:       {Sweetness = 10, Flavor < 8,                                                                Country = 'India'}
* James:      {Sweetness = 10, Flavor = 0,                        Aroma = 0,                              Country = 'United States (Hawaii)'}

I don't think Marcio needs to buy heaps of different coffee, maybe 1 bag per person

Marcio and Rafa, 2 bags
```{r Marcio and Rafa}
m_r <- copy(country_clusters[Country.of.Origin == "Brazil"])
m_r[, taste_distance := 
      abs(Sweetness - 10) +
      abs(Flavor - 8) +
      abs(Acidity - 8) +
      abs(Body - 8) +
      abs(Balance - 8) +
      abs(Aftertaste - 8)]
setorder(m_r, taste_distance)
m_r %>% head(2)
```

Marcio wanted to take Subu on a trip in Brazil, so find the Brazillian coffees
which are similar to Subu's taste. As they are going on a tour they can have 10
coffees?
```{r Subu Brazil}
subu_brazil <- copy(country_clusters[Country.of.Origin == "Brazil"])
subu_brazil[, taste_distance :=
              abs(Sweetness - 10) +
              Flavor
            ]
setorder(subu_brazil, taste_distance)
subu_brazil %>% head(10)
```

Subu India
```{r Subu India}
subu_india <- copy(country_clusters[Country.of.Origin == "India"])
subu_india[, taste_distance :=
             abs(Sweetness - 10) +
             Flavor]
setorder(subu_india, taste_distance)
subu_india %>% head(1)
```

James
```{r James}
james <- copy(country_clusters[Country.of.Origin == "United States (Hawaii)"])
james[, taste_distance :=
        abs(Sweetness - 10) +
        Flavor +
        Aroma]
setorder(james, taste_distance)
james %>% head(1)
```

